\LARGE \bf Planning with Attitude

### 2020-10-09 paper/main.tex
Clean up results section

    \subsection{Results}
    \begin{figure}
        \centering
        \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
        \caption{Convergence comparison for Wahbas problem. By performing Newton's method
        on the error quaternion and applying the result to the full quaternion we achieve
        quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
        error.}
    \end{figure}

\section{Trajectory Optimization on $\R^n \times SO(3)$}
    Here we outline the modifications to the ALTRO solver \cite{howell2019altro}, to
    solve trajectory optimization problems for rigid bodies, which extends easily to
    arbitrary systems in $R^n \times SO(3)$. ALTRO is an efficient solver for constrained
    nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented
    Lagrangian framework.

@@

        \subsubsection{Geodesic Distance}
            Alternatively, we can use the geodesic distance between two quaternions
            \cite{Kuffner2004},
            \begin{equation} \label{eq:quat_geodesic}
                J_\text{geo} = (1-\abs{q_g^T q}) ,
            \end{equation}
            whose gradient and Hessian are,
            \begin{align}
                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
            \end{align}
            where the sign of the Hessian corresponds to the sign of $q_d^T q$.

\section{Experiments} \label{sec:experiments}
    Code for all experiments is available on 
    \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
    {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.

@@

    \subsection{Trajectory Optimization}
        In this section we present several trajectory optimization problems solved using
        our modified version of ALTRO. The two quadrotor examples use 101 knot points for
        5 second trajectories, and the airplane example uses 51 knot points for a 1.25
        second trajectory. All results were run on a desktop computer with an Intel
        i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
        experiments are given in Fig. \ref{fig:timing_chart}.

        In all examples, MLQR is implemented with a unit quaternion attitude state and
        Rodrigues parameters as the error state. The first example compares the two cost
        functions given in Sec. \ref{sec:cost_functions}, while the last two use the
        geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
        \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
        from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
        vectors). All examples re-normalize the quaternion in the dynamics function and
        use a third-order explicit Runge-Kutta integrator.

@@

        \subsection{Airplane Barrel Roll}
        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
        tasked to do a barrel roll by setting a high terminal cost for being upside-down,
        see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
        conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
        converged faster than the pure quaternion version. Despite the extra matrix
        multiplications and extra Hessian term, it also was faster per iteration than its
        iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
        better performance by correctly leveraging the structure of the rotation group
        during the optimization routine.

@@

        \subsubsection{Quadrotor Flip}
        We sucessfully optimized a 360 degree flip trajectory for a quadrotor using the 
        modified version of ALTRO.
	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
	    to stay above the floor and move to a goal state 2 meters away in the $+y$
	    direction. The solver is initialized with a dynamically infeasible trajectory
	    that linearly interpolates between the initial and final states, rotating the
	    quad around the x-axis a full \ang{360}.

	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
	    ALTRO.  t

@@

\section{Conclusions} \label{sec:conclusion}
    We have presented a general, unified method for planning and control on rigid-body
    systems with arbitrary attitude using standard linear algebra and vector calculus. By
    applying these methods to LQR to correctly account for the group structure of
    rotations, we have matched or exceeded the performance of a state-of-the-art
    geometric controller designed specifically for quadrotors, while being more general,
    requiring less system-dependent tuning, having less computational overhead, and being
    easier to implement.

    We have also demonstrated a straight-forward way to adapt nonlinear trajectory
    optimization techniques to work with quaternion-valued states. For these problems, we
    found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
    computationally efficient and provide excellent convergence in practice. We recommend
    the use of unit quaternions within planning and control algorithms for their
    simplicity, computational efficiency, and lack of singularities. We further recommend
    the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.

    Future areas of research include the application of these methodologies to more
    complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
    more in-depth analysis of the convergence behavior of the algorithms proposed in the
    current work.

>>>>

    \subsection{Results}
    \begin{figure}
        \centering
        \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
        \caption{Convergence comparison for Wahbas problem. By performing Newton's method
        on the error quaternion and applying the result to the full quaternion we achieve
        quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
        error.}
        \label{fig:wahba_convergence}
    \end{figure}

\section{Trajectory Optimization on $\R^n \times SO(3)$} \label{sec:trajopt}
    Here we outline the modifications to the ALTRO solver \cite{howell2019altro}, to
    solve trajectory optimization problems for rigid bodies, which extends easily to
    arbitrary systems in $R^n \times SO(3)$. ALTRO is an efficient solver for constrained
    nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented
    Lagrangian framework.

@@

        \subsubsection{Geodesic Distance} \label{sec:geodesic}
            Alternatively, we can use the geodesic distance between two quaternions
            \cite{Kuffner2004},
            \begin{equation} \label{eq:quat_geodesic}
                J_\text{geo} = (1-\abs{q_g^T q}) ,
            \end{equation}
            whose gradient and Hessian are,
            \begin{align}
                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
            \end{align}
            where the sign of the Hessian corresponds to the sign of $q_d^T q$.

\section{Experiments} \label{sec:experiments}
    In this section we present several trajectory optimization problems for systems that
    undergo large changes in attitude: an airplane barrell roll, a quadrotor flip, and a
    satellite with flexible modes that must slew to a new orientation while avoiding a 
    ``keep out'' orientation zone. All results were run on a desktop computer with an 
    AMD Ryzen 2950x processor with 40 GB of RAM. 
    All problems are run using ALTRO, first without any of the modification presented 
    in the current paper, labeled ``original'', and then using the modifications listed
    in Sec. \ref{sec:trajopt} and the geodesic cost funtion described in Sec. \ref{sec:geodesic}.
    Timing results are summarized in Table \ref{tab:timing_results}.
    All experiments were solved to a constraint satisfaction tolerance of $10^{-5}$.
    Code for all experiments is available on 
    \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
    {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.

@@

@@

    \subsection{Airplane Barrel Roll}

        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
        tasked to do a barrel roll by setting a high terminal cost for being upside-down,
        see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
        conditions. The convergence of the different versions is compared in Fig.
        \ref{fig:c_max_convergence}. For both the quadrotor flip and the airplane barrel
        roll, the modified version of ALTRO converged faster than original version.
        For these highly aerobatic maneuvers, we achieve, as expected, better
        performance by correctly leveraging the structure of the rotation group during
        the optimization routine.

@@

    \subsection{Quadrotor Flip}
        We sucessfully optimized a 360 degree flip trajectory for a quadrotor using the 
        modified version of ALTRO.
	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
        of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular arc.
        The quadrotor was constrained
	    to stay above the floor and move to a goal state 2 meters away in the $+y$
	    direction. The solver was initialized with a dynamically infeasible trajectory
	    that linearly interpolates between the initial and final states, rotating the
        quad around the x-axis a full \ang{360}.

	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
        ALTRO. The original version of ALTRO, even after significant tuning efforts,
        would not converge to the desired solution. It instead got halfway and then ``unwound''
        360 degrees and then continued rotating to the final orientation. This behavior
        is common and expected when attempting optimization that does not properly account 
        for the grou structure of rotations.

@@

\section{Conclusions} \label{sec:conclusion}
    We have presented a general, unified method for planning and control on rigid-body
    systems with arbitrary attitude using standard linear algebra and vector calculus.
    We have demonstrated that the application of this methodology is straightforward and
    yields substantion improvements in the convergence of Newton-based methods (see Fig. 
    \ref{fig:wahba_convergence}) while also offering improvements for nonlinear constrained
    trajectory optimization for floating-base systems (see Table \ref{tab:timing_results}).

    With the modifications presented, ALTRO can solve problems few other methods for 
    trajectory optimization can. Many state-of-the-art methods such as direct 
    collocation or sequential convex programming rely on commercial, proprietary, or 
    general-purpose solvers whose internal numerics often abstracted away from the user. 
    By exploiting both the unique Markovian structure of the trajectory optimization problem
    and the group structure of rotations, ALTRO will likely be able to solve new classes of 
    problems with at or near real-time performance. Future work will focus on continued 
    refinement of the implementation and benchmarking against current state-of-the-art 
    methods for trajectory optimization.

    Additionally, the methods presented here can easily be leveraged to adapt other classes
    of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future
    work may include adaptation of methods for state estimation, localization, design, or 
    other methods for motion planning such as direct collocation.

    Future work will also include the application of these methods to multi-body robotic
    systems, such as humanoids or quadrupeds, especially in methods that exploit 
    ``maximal'' coordinates that include the 3D orientation of each body \cite{brudigam2020linear}.
